### 2016-11-26 dissertation/prelim_GF.tex
dissertation v2_part reacha update 1
dissertation v2_part normal complete

\chapter{Galois Field and Sequential Arithmetic Circuits}
This chapter provides a mathematical background for understanding 
finite fields (Galois fields) and explains how to design Galois field arithmetic circuits.
We first introduce the mathematical concepts of groups, rings, fields, and 
polynomials. 
We then apply these concepts to create Galois field arithmetic functions and 
explain how to map them to a Boolean circuit implementation.
Additonally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
as the normal basis theory behind the designing such hardware.

@@

If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
to binary bit-vector representations by recording the coefficients. For example,

\begin{table}[H]
\centering
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}
\begin{tabular}{|c|c||c|c|} 
\hline
$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
\hline
$0000$        & $0$           & $1000$  &$\alpha^3$\\
\hline
$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
\hline
$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
\hline
$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
\hline
$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\label{table:booltogalois}  
\end{table}

@@

\begin{Example}[NB multiplication using $\lambda$-Matrix]
Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
\begin{equation*}
M^{(0)} = \left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right).
\end{equation*}
i.e.,
\begin{equation*}
c_0 = (a_0\  a_1\  a_2)\left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)\left(
\begin{array} {lcr}
b_0\\
b_1\\
b_2
\end{array} \right)
\end{equation*}
From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
\begin{equation*}
M^{(1)} = \left(
\begin{array} {lcr}
1 & 0 & 1\\
0 & 0 & 1\\
1 & 1 & 0
\end{array} \right)~~~~~~
M^{(2)} = \left(
\begin{array} {lcr}
0 & 1 & 1\\
1 & 1 & 0\\
1 & 0 & 0
\end{array} \right)
\end{equation*}
\end{Example}

@@

\def\divrule#1#2{%
\noalign{\moveright#1\digitwidth%
\vbox{\hrule width#2\digitwidth}}}
111010111\,\begin{tabular}[b]{@{}r@{}}
101001 \\ \hline
\big)\begin{tabular}[t]{@{}l@{}}
11010110001110 \\
111010111 \\ \divrule{0}{14}
~~111101101 \\
~~111010111 \\ \divrule{2}{12}
~~~~~111010110 \\
~~~~~111010111 \\ \divrule{5}{9}
~~~~~~~~~~~~~1
\end{tabular}
\end{tabular}
\vspace{0.5cm}

@@

On the other hand, operands $A$ and $B$ can be written in NB as
$$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
is
(Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
\begin{equation*}
M^{(0)} = \left(\begin{array}{lccccccr}
0 &0 &0 &0 &1 &0 &1 &1 \\
0 &0 &1 &1 &1 &1 &0 &0 \\
0 &1 &0 &0 &0 &0 &1 &0 \\
0 &1 &0 &0 &1 &1 &0 &1 \\
1 &1 &0 &1 &0 &1 &0 &0 \\
0 &1 &0 &1 &1 &0 &0 &1 \\
1 &0 &1 &0 &0 &0 &0 &0 \\
1 &0 &0 &1 &0 &1 &0 &1
\end{array}\right)
\end{equation*}
Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
(or shifting $M^{(0)}$, either is applicable), we can successively
obtain other bits of product. The final answer is
$$C = (0000~0001)_{NB}$$
It is equivalent to the result in StdB.
\end{Example}

@@

\section{Design a Normal Basis Multiplier on Gate Level}
The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
routine of several kinds of modern NB multipliers.
\subsection{Sequential Multiplier with Parallel Outputs}
The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.

@@

\begin{Example}[5-bit Agnew's SMPO]
Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
$\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
The $0$-th $\lambda$-Matrix for this ONB is
\begin{equation*}
M^{(0)} = \left(\begin{array}{lcccr}
0 &1 &0 &0 &0 \\
1 &0 &0 &1 &0 \\
0 &0 &0 &1 &1 \\
0 &1 &1 &0 &0 \\
0 &0 &1 &0 &1
\end{array}\right)
\end{equation*}
Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.

@@

Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
\begin{equation}
\label{eqn:5bitSMPO_c0}
c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
\end{equation}
Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
remaining bits of the product, which can be written in following general form polynomial:
\begin{align}
c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
Note all index calculations are reduced modulo 5.

Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
In the first clock cycle,
signal sent to the flip-flop in block $R_0$ denotes function:
$$R_0^{(1)} = a_1b_0$$
It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
signal sent to flip-flop in block $R_1$ is:
$$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
$$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.

@@

Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
in figure \ref{fig:SMPO}. A brief proof is as follows:

@@

 \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
 \KwOut{$R=A\times B$}
%%%%%%%%%%%%%%%%%%%%
  Initialize $A,B$ and aux var $X$ to 0\;
  \For { ($i=0$;   $i < k$; ++i ) }
  {
	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
  }
  $R\gets X$
\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
\end{algorithm}

@@

To better understand the mechanism of RH-SMPO, we will use this 5-bit 
RH-SMPO as an example and introduce the details on how to design it.
\begin{Example}[Designing a 5-bit RH-SMPO]
From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
describes, we implement auxiliary function $F_{k-1}$ in the logic:
$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
\begin{equation}
\label{eqn:5bitRHaux}
F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
\end{equation}
Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.

@@

It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
$$M_{i,j}^{(0)} = T_{j-i,-i}$$
with indices reduced modulo $k$ (proof given in appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:

@@

Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
$e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
\begin{equation*}
\beta\cdot\beta^2 = 
\begin{bmatrix}
1 & 0 & 0 & 1 & 0
\end{bmatrix}
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
\end{bmatrix}
= \beta + \beta^{2^3}
\end{equation*}

@@

Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
After gathering all interconnection information, we can translate it to gate-level circuit implementation:
$$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$

Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
$R_{i-1}$ in last clock cycle, we can connect using relation 
$$R_i = r_{i-1} + e_i$$
In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.

>>>>

\chapter{Galois Field and Sequential Arithmetic Circuits}
\label{ch:prelim_GF}
This chapter provides a mathematical background for understanding 
finite fields (Galois fields) and explains how to design Galois field arithmetic circuits.
We first introduce the mathematical concepts of groups, rings, fields, and 
polynomials. 
We then apply these concepts to create Galois field arithmetic functions and 
explain how to map them to a Boolean circuit implementation.
Additonally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
as the normal basis theory behind the designing such hardware.

@@

If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
to binary bit-vector representations by recording the coefficients. For example, for elements in
$\F_{2^4}$, the results are shown in table \ref{tab:gfelement}, column ``Polynomial".

% \begin{table}[H]
% \centering
% \caption{Bit-vector, Exponential and Polynomial representation of
% elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
% \pmod{x^4+x^3+1}$}
% \begin{tabular}{|c|c||c|c|} 
% \hline
% $a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
% \hline
% $0000$        & $0$           & $1000$  &$\alpha^3$\\
% \hline
% $0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
% \hline
% $0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
% \hline
% $0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
% \hline
% $0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
% \hline
% $0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
% \hline
% $0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
% \hline
% $0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
% \hline
% \end{tabular}
% \label{table:booltogalois}  
% \end{table}

@@

\begin{Example}[NB multiplication using $\lambda$-Matrix]
Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
\begin{equation*}
M^{(0)} = \left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)
\end{equation*}
i.e.,
\begin{equation*}
c_0 = (a_0\  a_1\  a_2)\left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)\left(
\begin{array} {lcr}
b_0\\
b_1\\
b_2
\end{array} \right)
\end{equation*}
From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
\begin{equation*}
M^{(1)} = \left(
\begin{array} {lcr}
1 & 0 & 1\\
0 & 0 & 1\\
1 & 1 & 0
\end{array} \right)~~~~~~
M^{(2)} = \left(
\begin{array} {lcr}
0 & 1 & 1\\
1 & 1 & 0\\
1 & 0 & 0
\end{array} \right)
\end{equation*}
\end{Example}

@@

\begin{center}
\def\divrule#1#2{%
\noalign{\moveright#1\digitwidth%
\vbox{\hrule width#2\digitwidth}}}
111010111\,\begin{tabular}[b]{@{}r@{}}
101001 \\ \hline
\big)\begin{tabular}[t]{@{}l@{}}
11010110001110 \\
111010111 \\ \divrule{0}{14}
~~111101101 \\
~~111010111 \\ \divrule{2}{12}
~~~~~111010110 \\
~~~~~111010111 \\ \divrule{5}{9}
~~~~~~~~~~~~~1
\end{tabular}
\end{tabular}
\end{center}
\vspace{0.5cm}

@@

On the other hand, operands $A$ and $B$ can be written in NB as
$$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
is
(Computation of $\lambda$-Matrix refers to Appendix \ref{append:NB})
\begin{equation*}
M^{(0)} = \left(\begin{array}{lccccccr}
0 &0 &0 &0 &1 &0 &1 &1 \\
0 &0 &1 &1 &1 &1 &0 &0 \\
0 &1 &0 &0 &0 &0 &1 &0 \\
0 &1 &0 &0 &1 &1 &0 &1 \\
1 &1 &0 &1 &0 &1 &0 &0 \\
0 &1 &0 &1 &1 &0 &0 &1 \\
1 &0 &1 &0 &0 &0 &0 &0 \\
1 &0 &0 &1 &0 &1 &0 &1
\end{array}\right)
\end{equation*}
Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
(or shifting $M^{(0)}$, either is applicable), we can successively
obtain other bits of product. The final answer is
$$C = (0000~0001)_{NB}$$
It is equivalent to the result in StdB.
\end{Example}

@@

\section{Design a Normal Basis Multiplier on Gate Level}
\label{sec:nbdesign}
The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
routine of several kinds of modern NB multipliers.
\subsection{Sequential Multiplier with Parallel Outputs}
The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.

@@

\begin{Example}[5-bit Agnew's SMPO]
Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
$\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
The $0$-th $\lambda$-Matrix for this ONB is
\begin{equation*}
M^{(0)} = \left(\begin{array}{lcccr}
0 &1 &0 &0 &0 \\
1 &0 &0 &1 &0 \\
0 &0 &0 &1 &1 \\
0 &1 &1 &0 &0 \\
0 &0 &1 &0 &1
\end{array}\right)
\end{equation*}
Then a typical design of 5-bit Agnew's SMPO is depicted in Figure \ref{fig:SMPO}.

@@

Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
\begin{equation}
\label{eqn:5bitSMPO_c0}
c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
\end{equation}
Note it is written in the form of Equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
remaining bits of the product, which can be written in following general form polynomial:
\begin{align}
c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
Note all index calculations are reduced modulo 5.

Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
In the first clock cycle,
signal sent to the flip-flop in block $R_0$ denotes function:
$$R_0^{(1)} = a_1b_0$$
It equals to the first term of Equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
signal sent to flip-flop in block $R_1$ is:
$$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
which forms first 2 terms of Equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
$$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
which equals to $c_0$ in Equation \ref{eqn:5bitSMPO_c0}.
After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.

@@

Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
in Figure \ref{fig:SMPO}. A brief proof is as follows:

@@

 \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
 \KwOut{$R=A\times B$}
%%%%%%%%%%%%%%%%%%%%
  Initialize $A,B$ and aux var $X$ to 0\;
  \For { ($i=0$;   $i < k$; ++i ) }
  {
	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Equation \ref{eqn:aux}*/\;}
	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
  }
  $R\gets X$
\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
\end{algorithm}

@@

To better understand the mechanism of RH-SMPO, we will use this 5-bit 
RH-SMPO as an example and introduce the details on how to design it.
\begin{Example}[Designing a 5-bit RH-SMPO]
From Equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
and XOR gates in $c$-layer according to Equation \ref{eqn:auxDC}. Concretely, as Algorithm \ref{alg:RHmulti}
describes, we implement auxiliary function $F_{k-1}$ in the logic:
$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
\begin{equation}
\label{eqn:5bitRHaux}
F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
\end{equation}
Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.

@@

It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
$$M_{i,j}^{(0)} = T_{j-i,-i}$$
with indices reduced modulo $k$ (proof given in Appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:

@@

Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
$e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in Equation \ref{eqn:multitable}:
\begin{equation*}
\beta\cdot\beta^2 = 
\begin{bmatrix}
1 & 0 & 0 & 1 & 0
\end{bmatrix}
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
\end{bmatrix}
= \beta + \beta^{2^3}
\end{equation*}

@@

Let us look back at Equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
After gathering all interconnection information, we can translate it to gate-level circuit implementation:
$$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$

Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
$R_{i-1}$ in last clock cycle, we can connect using relation 
$$R_i = r_{i-1} + e_i$$
In this example, according to the M-table in Figure \ref{fig:multitable}, columns $e_1,e_2$
have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.

### 2016-11-26 dissertation/prelim_basis.tex
dissertation v2_part normal complete

{\bf Problem setup:} Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of the Galois field arithmetic circuit over $\Fkk$, as figure \ref{fig:abs_setup} shows.
This circuit performs some unknown function $f$ over 
$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
output and $A_1,\dots,A_n$ are the $k$-bit inputs.
The polynomial representation of $\Func$ over $\Fkk$ is thus:

@@

\begin{Example}
\label{exp:finalMul2Bit}
Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
$P(x)=x^{2}+x+1$, given in Fig. \ref{fig:2bitmul}. Variables $a_0,
a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and
$c_0, c_1, c_2, c_3, r_0$ are intermediate variables. 
%The gate $\otimes$
%corresponds to AND-gate, i.e. bit-level multiplication modulo 2. The
%gate $\oplus$ corresponds to XOR-gate, i.e. addition modulo 2.

>>>>

{\bf Problem setup:} Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of the Galois field arithmetic circuit over $\Fkk$, as Figure \ref{fig:abs_setup} shows.
This circuit performs some unknown function $f$ over 
$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
output and $A_1,\dots,A_n$ are the $k$-bit inputs.
The polynomial representation of $\Func$ over $\Fkk$ is thus:

@@

\begin{Example}
\label{exp:finalMul2Bit}
Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
$P(x)=x^{2}+x+1$, given in Figure \ref{fig:2bitmul}. Variables $a_0,
a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and
$c_0, c_1, c_2, c_3, r_0$ are intermediate variables. 
%The gate $\otimes$
%corresponds to AND-gate, i.e. bit-level multiplication modulo 2. The
%gate $\oplus$ corresponds to XOR-gate, i.e. addition modulo 2.
